---
title: "rscropt"
output: html_document
date: "2025-12-15"
---

## Cumulative Link Mixed Effect Model

Install the following packages for analysis

```{r}

install.packages("tidyverse")
library(tidyverse)
install.packages("ordinal")
library(ordinal)
install.packages("MASS")
library(MASS)
install.packages("effects")
library(effects)
library(emmeans)
```

Load the data for CLMM

```{r}
d1 <- read.csv('../data/d1.csv') #the full path depends on where you saved d1. 
d1 <- as.tibble(d1)
```

Code each row with their appropriate types of variable (character, factor, ordered factor, etc.)

```{r}
d1$subject <- as.character(d1$subject)
d1$item <- as.character(d1$item)
d1$Rating <- as.ordered(d1$Rating)
d1$Trans <- as.numeric(d1$Trans)
d1$Anim <- as.factor(d1$Anim)
```

Fit the data using CLMM

```{r}
#Three models for comparison
N_modeld1 <- clmm(Rating~1+(1+subject)+(0+Anim|subject)+(0+Trans|subject)(1|item), data=d1, Hess=TRUE)
A_modeld1 <- clmm(Rating~1+Anim+(1+subject)+(0+Anim|subject)+(0+Trans|subject)(1|item), data=d1, Hess=TRUE)
AT_modeld1 <- clmm(Rating~1+Anim+Trans+(1+subject)+(0+Anim|subject)+(0+Trans|subject)(1|item), data=d1, Hess=TRUE)

```

Create summary statistics tables and figures

```{r}
#Summary tables for each model
summary(N_modeld1) #Null model
summary(A_modeld1) #A model
summary(AT_modeld1) #AT model

#anova table (for model selection)
anova(N_modeld1, A_modeld1, AT_modeld1)

#Effect plot (using effects package)
eff.1 <- Effect(focal.predictors = c("Anim"), AT_modeld1)
plot(eff.1, rug = FALSE)
eff.2 <- Effect(focal.predictors = c("Trans"), AT_modeld1)
plot(eff.2, rug = FALSE)
eff.all <- Effect(focal.predictors = c("Anim", "Trans"), AT_modeld1)
plot(eff.all, rug = FALSE)

```

## Cumulative Link Mixed Effect Model for d2 (factor of interest: causation)

Install the following packages for analysis

```{r}

install.packages("tidyverse")
library(tidyverse)
install.packages("ordinal")
library(ordinal)
install.packages("MASS")
library(MASS)
install.packages("effects")
library(effects)
library(emmeans)
```

Load the data for CLMM

```{r}
d2 <- read.csv('../data/d2.csv')
d2 <- as.tibble(d2)
d2
```

Code each row with their appropriate types of variable (character, factor, ordered factor, etc.)

```{r}
d2$subject <- as.character(d2$subject)
d2$item <- as.character(d2$item)
d2$Rating <- as.ordered(d2$Rating)
d2$Caus <- as.factor(d2$Caus)
d2$Trans <- as.ordered(d2$Trans)
```

#Relevel so Poss is the reference level
d2$Caus <- relevel(factor(d2$Caus), ref = "Poss")

Fit the data using CLMM

```{r}
#Two models for comparison
N_modeld2 <- clmm(Rating~1+(1+Caus|subject)+(1|item), data=d2, Hess=TRUE)
C_modeld2 <- clmm(Rating~1+Caus+(1+Caus|subject)+(1|item), data=d2, Hess=TRUE)
```

Create summary statistics tables and figures

```{r}
#Summary tables for each model
summary(N_modeld2) #Null model
summary(C_modeld2) #C model

#anova table (for model selection)
anova(N_modeld2, C_modeld2)

#pairwise comparison for C model (using emmeans package)
pw <- emmeans(C_modeld2, list(pairwise ~ Caus), adjust="bonferroni")

#Effect plot (using effects package)
eff.1 <- Effect(focal.predictors = c("Caus"), C_modeld2)
plot(eff.1, rug = FALSE)

```

## Descriptive statistics

### Plot the descriptive summary of d1

```{r}
#Create a list of items that corresponds to the labels for the bar plot
d1exe <- c("1飛んできた矢が的を射抜いた", "2ミサイルがそのビルを破壊した", "3飛んできたボールが窓ガラスを割った", "4落石が歩行者を押し潰した", "5一発の銃弾が兵士を殺した。", "6落雷が彼を気絶させた", "7山火事の炎が木々を焼き尽くした", "8この爆弾は装甲の厚い戦車を破壊できる", "9爆撃機から落とされる爆弾が街のあちこちを破壊している",
           "10雪崩がたくさんの登山客たちを飲み込んだ", "11急な土砂降りの雨が帰宅する人たちをずぶ濡れにした", "12強烈な突風が歩行者たちを転倒させた", "13津波が村を襲った", "14突風が看板を動かした", "15釘が自転車をパンクさせた", "16その危険な技が彼を骨折させた",
           "17麻薬が彼を殺した", "18壊れた遊具がその子供を怪我させた", "19強い風が庭の木を揺らしている", "20大地震が東京を襲うかもしれない", "21打球が観客席を超えて行った", "22サーチライトの光が脱獄囚を追っている", "23有毒ガスがその兵士を嘔吐させた", "24落ちてきた看板が歩行者を直撃した", "25凍結した路面がトラックをスリップさせた", "26この薬品はプラスチックを溶かす",
           "27雲が太陽を隠した", "28工場の煙がその地区の住民たちを困らせた", "29不気味な絵画が美術館に来た子供たちを怖がらせた", "30監視カメラが逃げる犯人を捉えた", "31紛れ込んだ部品がベルトコンベアを止めた", "32満月が家々を照らしている", "33バラの棘が指を刺した", "34重い年貢がその村人たちを苦しめた", "35防弾チョッキが彼を守った", "36長い不況が国民を苦しめた", "37紅葉が山を覆い始めた", "38このロープは重い展示物を支えることができる", "39大きな川が2つの村を隔てている", "40初戦の勝利が選手たちを鼓舞した", "41この新作映画は観客を魅了するだろう", "42この保険は高齢者を対象にしている", "43大きな木が枝を揺らしている", "44このエレベーターは緊急停止装置を備えている", "45その薬はいろいろな効能を持つ", "46この法律における「学生」の定義は専門学校生を含む", "47政治は変えられないという思い込みが国民を支配している", "48祭りの季節になり、熱気が人々を包みはじめた")

#Bar plot using ggplot2 

d1bar <- ggplot(d1)+geom_bar(aes(y = as.factor(item), fill = ordered(Rating, rev(levels(Rating)))), position = position_fill())+scale_x_continuous(labels = scales::percent)+
  scale_y_discrete(limit=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", "21", "22", "23", "24", "25", "26", "27", "28", "29", "30", "31", "32", "33", "34", "35", "36", "37", "38", "39", "40", "41", "42", "43", "44", "45", "46", "47", "48"), labels=d1exe, theme_set( theme_bw(base_family = "HiraKakuProN-W3")))+labs(x = "Percentage")+theme_set( theme_bw(base_family = "HiraKakuProN-W3"))

print(d1bar)
```

### Plot the descriptive summary of d2

```{r}
#Create a list of items that corresponds to the labels for the bar plot

d2txt <- c("1地震が建物を倒壊させた", "2急な段差が視覚障害の男性をつまずかせた", "3突風が看板を倒した", "4鉄板の熱がバターを溶かした", "5長い年月をかけて風が砂丘を作った", "6雨粒が窓ガラスを叩いた","7燃え盛る炎が消防士を火傷させた", "8乱気流がキャビンアテンダントをよろめかせた", "9鍵がドアを開けた", "10バンドエイドが擦り傷を隠した", "11チェーンソーが木を切った", "12電子レンジが惣菜を温めた", "13ハンマーが窓ガラスを割った", "14赤ペンがバツ印をつけた", "15エクセルがデータをまとめた", "16懐中電灯が草むらを照らした", "17美しいダンスが観客を魅了した", "18借金が彼を苦しめた", "19激しい雷鳴が子供達を怖がらせた", "20干ばつが村人たちを悩ませた", "21久しぶりの雨が村人を喜ばせた", "22高速道路の騒音が住民たちを困らせている", "23その大きな犬が子供たちを怯えさせた", "24見事なマジックが観客を驚かせた", "25倒れた木が道路を塞いでいる", "26マングローブがその土地全体を覆っている", "27珊瑚礁が島を囲んでいる", "28川が市の東側を流れている", "29山地が国土の大半を占めている", "30物価上昇率が賃金上昇率を上回っている", "31この避難所はたくさんの被災者を収容できる", "32高い壁が西側と東側の人々を隔てている", "33このヒーターは自動オフ機能を備えている", "34庭の木が立派な実をつけている", "35ここの井戸水は豊富なミネラルを含んでいる", "36この薬は強い副作用を持つ", "37今度のアルバムは新曲をたくさん収録している", "38このウォークイン・クローゼットはたくさんの服を収納できる", "39この国は大量の失業者を抱えている", "40この単語は文脈によっていろいろな意味を持つ")

#Bar plot using ggplot2 

d2bar <- ggplot(d2)+geom_bar(aes(y = as.factor(item), fill = ordered(Rating, rev(levels(Rating)))), position = position_fill())+scale_x_continuous(labels = scales::percent)+ scale_y_discrete(limit=c("1", "2","3", "4", "5", "6", "7","8","9", "10", "11", "12", "13","14", "15","16", "17", "18","19", "20","21", "22", "23","24", "25", "26","27", "28", "29","30", "31","32", "33", "34", "35", "36", "37", "38", "39", "40"), labels=d2txt, theme_set( theme_bw(base_family = "HiraKakuProN-W3")))+labs(x = "Percentage")+theme_set( theme_bw(base_family = "HiraKakuProN-W3")) 

print(d2bar)
```
